Semantic Textual Similarity using deep learning

### 2017-06-05 paper/main.tex
Part of deep learning explanation added

\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}

@@

\begin{table}[h!]
\caption{STS score description, adapted from~\citep{agirre2016semeval}}
\label{tab:sts-score}
\begin{center}
\begin{tabularx}{0.45\textwidth}{llr}
\toprule
Score & Explanation\\
\midrule
5 & \textit{Two sentences are completely equivalent}\\
& The bird is bathing in the sink\\
& Birdie is washing itself in the water basin.\\
\midrule
4 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
  & \multicolumn{1}{m{6cm}}{In May 2010, the troops attempted to invade
Kabul.} \\
& \multicolumn{1}{m{6cm}}{The US army invaded Kabul on May 7th last
year, 2010.} \\
\midrule
3 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
  & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
suspect.}\\
& ``He is not a suspect anymore.'' John said.\\
\midrule
2 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
&They flew out of the nest in groups. \\
&They flew into the nest together. \\
\midrule
1 & \multicolumn{1}{m{6cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
& The woman is playing the violin.\\
& The young lady enjoys listening to the guitar.\\
\midrule
0 & \textit{The two sentences are completely dissimilar.}\\
  & \multicolumn{1}{m{6cm}}{John went horse back riding at dawn with a whole group of friends.}\\
  & \multicolumn{1}{m{6cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
\bottomrule
\end{tabularx}
\end{center}
\end{table*}

@@

One of the simplest approaches to measuring semantic similarity is the Jaccard rindex.
It is a statistic used for comparing similarity and diversity of finite sets.
Let $J(S_1, S_2)$ represent the Jaccard index between two sentences $S_1$ and $S_2$.
The similariy between sentences is defined as:
\begin{equation}
	\label{eq:jaccard}
	J(S_1, S_2) = \frac{|S_1 \cap S_2|}{|S_1 \cup S_2|},
\end{equation}

@@

Simple approach to extending the word embedding system to a sentence is simply performing some aggregation operation on each of the embeddings.
We used the summation function aggregation on the words of the sentence, after removing stopwords.
Stopwords list is used from the NLTK library.
Word embedding used is the English version of the FastText word embeddings \citep{joulin2016fasttext}.

@@

Deep neural networks have been shown to perform well on language modelling tasks
(insert citation). Use of highly non-linear functions in the models allows for learning of information on a high level of abstraction. Semantic textual similarity bla bla idem spavat sutra ovo dovrsim.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{listings}
\lstset{
   breaklines=true,
   basicstyle=\ttfamily}

@@

\begin{table}[h!]
\caption{STS score description, adapted from~\citep{agirre2016semeval}}
\label{tab:sts-score}
\begin{center}
	\resizebox{\columnwidth}{!}{
\begin{tabularx}{\columnwidth}{llr}
\toprule
Score & Explanation\\
\midrule
5 & \textit{Two sentences are completely equivalent}\\
& The bird is bathing in the sink\\
& Birdie is washing itself in the water basin.\\
\midrule
4 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are mostly equivalent, but some unimportant details differ}} \\
  & \multicolumn{1}{m{6.5cm}}{In May 2010, the troops attempted to invade
Kabul.} \\
& \multicolumn{1}{m{6.5cm}}{The US army invaded Kabul on May 7th last
year, 2010.} \\
\midrule
3 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are roughly equivalent, but some important information differs/missing.}} \\
  & \multicolumn{1}{m{6cm}}{John said he is considered a witness but not a
suspect.}\\
& ``He is not a suspect anymore.'' John said.\\
\midrule
2 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but share some details.}} \\
&They flew out of the nest in groups. \\
&They flew into the nest together. \\
\midrule
1 & \multicolumn{1}{m{6.5cm}}{\textit{The two sentences are not equivalent, but are on the same topic.}}\\
& The woman is playing the violin.\\
& The young lady enjoys listening to the guitar.\\
\midrule
0 & \textit{The two sentences are completely dissimilar.}\\
  & \multicolumn{1}{m{6.5cm}}{John went horse back riding at dawn with a whole group of friends.}\\
  & \multicolumn{1}{m{6.5cm}}{Sunrise at dawn is a magnificent view to take in if you wake up early enough for it.}\\
\bottomrule
\end{tabularx}
}
\end{center}
\end{table*}

@@

One of the simplest approaches to measuring semantic similarity is the Jaccard index.
It is a statistic used for comparing similarity and diversity of finite sets.
Let $J(S_1, S_2)$ represent the Jaccard index between two sentences $S_1$ and $S_2$.
The similariy between sentences is defined as:
\begin{equation}
	\label{eq:jaccard}
	J(S_1, S_2) = \frac{|S_1 \cap S_2|}{|S_1 \cup S_2|},
\end{equation}

@@

Simple approach to extending the word embedding system to a sentence is simply performing some aggregation operation on each of the embeddings and then performing the cosine similarity measure.
We used the summation function aggregation on the words of the sentence, after removing stopwords.
Stopwords list is used from the NLTK library.
Word embedding used is the English version of the FastText word embeddings \citep{joulin2016fasttext}.

@@

\subsubsection{Motivation}
Although feature engineering measures are able to provide an approximate measure of similarity, it is trivial to reverse-engineer adversarial examples that completely break the system.
Consider, for example, two following sentences:

\begin{lstlisting}
Compared to me, he's heavier.
I have less kilograms than him.
\end{lstlisting}

Although their Jaccard index is zero (they share no common words), their semantic similarity is high.

Similar examples can easily be generated by replacing any specific word with its definition:
\begin{lstlisting}
Backpropagation.
Reverse-mode automatic differentiation.
\end{lstlisting}

Concepts on a such high level of abstraction easily break even the strongest systems since new word definitions could be defined arbitrarily.
No known models are able to generalize well to strong adversarial examples in an arbitrary language. 
Such high level generalization could be considered as actual \textit{understanding} of the natural language, which is an AI-complete problem.

A step forward to a model which could provide strong generalization capabilities is the one that has little to none domain specific knowledge embedded inside it and one which has no feature engineering.

One of such models are deep neural networks.

In this paper, one specific instance of such a model is explored: character-level language modelling with deep recurrent neural networks.

\subsubsection{LSTM networks}

Other than human brains, No known models are able to generalize well to such

Although word embeddings are, in theory, more robust to simple word changes,

Deep neural networks have been shown to perform well on language modelling tasks
(insert citation). 
The non-linear nature of neural networks allows them to model concepts on a high level of abstraction.

Semantic textual similarity exactly that: modelling similarity between concepts that are not shallow, like numbe
